---
title: "Crime Rate and HIV in NYC"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---


```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(plotly)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .6,
  fig.height = 8,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```




```{r, message=FALSE, warning=FALSE}
total_crime_df = 
  read_csv("./data/total_crime.csv")

census_df = 
  read_csv("./data/census.csv")

hiv_df = 
  read_csv("./data/hiv_complete.csv")

hiv_df1 =
  hiv_df %>% 
  filter(sex == "All",
         sex_cat == "All",
         race_cat == "All",
         age_cat == "All",
         risk_cat == "All",
         borough == "All") %>%
  select(year, borough_cat, total_hiv_diag_N) %>% 
  filter(!borough_cat %in% c("All", "Unknown borough", "Outside NYC")) %>% 
  rename(borough = borough_cat)

hiv_crime =
  left_join(hiv_df1, total_crime_df, by = c("year", "borough"))

hiv_crime = 
  hiv_crime %>% 
  left_join(census_df, by = "borough")

hiv_crime = 
  hiv_crime %>% 
  mutate(
    hiv_diag_rate = (total_hiv_diag_N / population_2010) * 10000,
    crime_rate = (total_crime / population_2010) * 10000,
    year = as.integer(year))
  
```


## Crime Rate over Years
```{r, message=FALSE, warning=FALSE}
crime_plot = 
  hiv_crime %>% 
  ggplot(aes(x = year, y = crime_rate, color = borough)) +
  geom_point(alpha = 0.7) +
  geom_line(size = 0.5, alpha = 0.5) +
  scale_x_continuous(
    breaks = c(2013, 2014, 2015, 2016, 2017, 2018, 2019),
    labels = c("2013", "2014", "2015", "2016", "2017", "2018", "2019"),
    limits = c(2013, 2019)) +
  labs(title = "Crime Rate Over Years In Each Borough",
       x = "Year",
       y = "Crime Rate (case/10k population)") 

ggplotly(crime_plot) %>% 
  layout(legend = list(orientation = "h", 
                       x = 0.1, y = -0.3))
```

**_Comments_**

The unemployment rate decreases over time from 2015 to 2019 for Bronx (Bronx County), Brooklyn (Kings County), Manhattan (New York County), Queens (Queens County), Staten Island (Richmond County).




## HIV Diagnosis Rate VS. Crime Rate in NYC

```{r, message=FALSE, warning=FALSE}
hiv_crime_plot =
  hiv_crime %>% 
  mutate(year = as.factor(year)) %>% 
  ggplot(aes(x = crime_rate, y = hiv_diag_rate)) +
  geom_point(aes(shape = borough, color = year), alpha = 0.8) +
  geom_smooth(size = 0.5) +
  labs(title = "HIV Diagnosis Rate VS. Crime Rate in NYC",
       x = "Crime Rate (case/10k population)",
       y = "HIV Diagnosis Rate (case/10k population)")

ggplotly(hiv_crime_plot)
```

*NOTE: the crime number only involves 7 major felony offenses according to NYPD: murder and non-negligent manslaughter, rape, robbery, felony assault, burglary, grand larceny and grand larceny of a vehicle.*

**_Comments_**

There is a positive correlation between crime rate and HIV diagnosis rate. In Manhattan, HIV diagnosis declined from 2013 to 2019, but crime rate did not change much.