---
title: "HIV Overview"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(plotly)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r, message=FALSE, warning=FALSE}
hiv_data = read_csv('./data/hiv_complete.csv')

hiv_diagnoses = 
  hiv_data %>% 
  filter(borough == "All", sex == "All") %>% 
  filter(borough_cat == "All", sex_cat == "All", race_cat == "All", age_cat == "All", risk_cat == "All") %>% 
  select(year, total_hiv_diag_N, death_N, plwha_N, borough_cat) %>% 
  pivot_longer(
    total_hiv_diag_N:death_N, 
    names_to = "outcome_cat",
    names_prefix = "_N",
    values_to = "outcome_number"
  ) %>% 
   mutate(
    outcome_cat = recode(outcome_cat, "total_hiv_diag_N" = "HIV diagnoses", "death_N" = "HIV Related Death")
    )
```


## HIV Trend Over Years (2011-2019)
```{r, message=FALSE, warning=FALSE}
hiv_diagnoses_plot = 
  hiv_diagnoses %>% 
  ggplot(aes(x = year, y = outcome_number, color = outcome_cat)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  scale_x_continuous(
    breaks = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019),
    labels = c("2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019")
  ) +
  scale_y_continuous(
    breaks = c(0, 500, 1000, 1500, 2000, 2500, 3000, 3500),
    labels = c("0", "500", "1000", "1500", "2000", "2500", "3000", "3500")
  ) +
  theme(axis.text.x = element_text(angle = 25, vjust = 0.5, hjust = 0.5)) +
  labs(title = "Trend Over Years",
       x = "Year",
       y = "Number") 

ggplotly(hiv_diagnoses_plot) %>% 
  layout(legend = list(orientation = "h", 
                       x = 0.1, y = -0.3))
```

**_Comments_**

According to this overall trend, the HIV diagnoses and HIV related death decreased over the years. 

```{r, message=FALSE, warning=FALSE}
hiv = 
  hiv_data %>% 
  filter(borough != "All") %>% 
  filter(borough_cat == "All", sex_cat == "All", race_cat == "All", age_cat == "All", risk_cat == "All") %>% 
  select(year, total_hiv_diag_N, borough, sex) %>% 
  group_by(borough, year) %>% 
  summarise(num_hiv_diagnoses = sum(total_hiv_diag_N))
```


## HIV Diagnosis by Boroughs
```{r, message=FALSE, warning=FALSE}
hiv_plot = 
  hiv %>% 
  ggplot(aes(x = year, y = num_hiv_diagnoses, fill = year)) +
  geom_bar(width = 0.5, stat = "identity", show.legend = T) +
  facet_grid(.~borough) +
  scale_x_continuous(
    breaks = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019),
    labels = c("2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019")
  ) +
  scale_y_continuous(
    breaks = c(0, 200, 400, 600, 800, 1000),
    labels = c("0", "200", "400", "600", "800", "1000")
  ) +
  theme(axis.text.x = element_text(angle = 75, vjust = 1, hjust = 0.5)) +
  labs(title = "Number of HIV Diagnosis",
       x = NULL, y = NULL) 

ggplotly(hiv_plot) %>% 
  layout(legend = list(orientation = "h", 
                       x = 0.1, y = -0.3))
```

**_Comments_**

Overall, Staten island has the lowest number of HIV diagnoses for each year from 2011 to 2017. The trend of HIV diagnoses for Bronx, Brooklyn, Manhattan, and Queens are similar to each other with a overall decreasing trend over years. 


```{r, message=FALSE, warning=FALSE}

hiv_risk_df =
  hiv_data %>%
  filter(borough != "All", risk_cat != "All") %>%
  select(year, borough, risk_cat, total_hiv_diag_N) %>% 
  group_by(borough, risk_cat) %>% 
  summarize(hiv_diag_over_years = sum(total_hiv_diag_N))

hiv_risk_total =
  hiv_risk_df %>% 
  group_by(borough) %>% 
  summarize(hiv_total = sum(hiv_diag_over_years))

hiv_risk_df1 = 
  hiv_risk_df %>% 
  left_join(hiv_risk_total, by = "borough") %>% 
  mutate(
    percentage = hiv_diag_over_years / hiv_total,
    risk_cat = fct_reorder(risk_cat, percentage)) %>% 
  rename(risk_factor = risk_cat)
```


## HIV Risk by Borough

```{r, message=FALSE, warning=FALSE}

hiv_risk_plot =
  hiv_risk_df1 %>% 
  ggplot(aes(x = risk_factor, y = percentage, fill = borough)) +
      geom_bar(width = 0.5, stat = 'identity', show.legend = F) +
      coord_flip() +
      facet_grid(.~borough) +
      labs(title = 'HIV Risk by Borough',
           x = NULL,
           y = NULL) + 
  theme(axis.text.x = element_text(size = 6, angle = 45),
        legend.title = element_blank())

ggplotly(hiv_risk_plot)  %>% 
  layout(legend = list(orientation = "h",
                     x = 0.1, y = -0.1))

  
```

**_Comments_**

The most common risk factor overall are men who have sex with men and heterosexual contact, which account for about 70% of HIV risk for all five boroughs. Sexual transmission is the main HIV transmission path in NYC.