---
title: "Unemployment in NYC"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(plotly)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r, message=FALSE, warning=FALSE}
unemployment_df = read_csv('./data/unemployment.csv')

unemployment_rate = 
  unemployment_df %>% 
  filter(month == "Avg") %>% 
  select(unemployment_rate, area, year, month) %>% 
  mutate(
    area = recode(area, "Bronx County" = "Bronx", "Kings County" = "Brooklyn", "New York  County" = "Manhattan", "Queens County" = "Queens", "Richmond County" = "Staten Island")
    )
```


## Unemployment Distribution Over Years
```{r, message=FALSE, warning=FALSE}
unemployment_plot = 
  unemployment_rate %>% 
  ggplot(aes(x = year, y = unemployment_rate, color = area)) +
  geom_point(alpha = 0.7) +
  geom_smooth(size = 0.5, alpha = 0.5) +
  theme(axis.text.x = element_text(angle = 25, vjust = 0.5, hjust = 0.5)) +
  labs(title = "Unemployment Rate Within Each Borough",
       x = "Year",
       y = "Unemployment Rate (%)") 

ggplotly(unemployment_plot) %>% 
  layout(legend = list(orientation = "h", 
                       x = 0.1, y = -0.3))
```

**_Comments_**

The unemployment rate decreases over time from 2015 to 2019 for Bronx (Bronx County), Brooklyn (Kings County), Manhattan (New York County), Queens (Queens County), Staten Island (Richmond County).




```{r, message=FALSE, warning=FALSE}
hiv_data = read_csv('./data/hiv_complete.csv')

hiv_diagnoses = 
  hiv_data %>% 
  filter(borough == "All", sex == "All") %>% 
  filter(borough_cat != "All") %>% 
  select(year, total_hiv_diag_N, borough_cat) %>% 
  filter(borough_cat %in% c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island")) %>% 
  filter(year %in% c("2015", "2016", "2017", "2018", "2019")) %>% 
  rename(area = borough_cat)

hiv_unemployment = 
  merge(hiv_diagnoses, unemployment_rate)
```

## Unemployment Rate VS. HIV Diagnoses 
```{r, message=FALSE, warning=FALSE}
hiv_unemployment_plot = 
  hiv_unemployment %>% 
  ggplot(aes(x = unemployment_rate, y = total_hiv_diag_N)) +
  geom_point(aes(shape = area, color = year)) +
  geom_smooth(size = 0.3) +
  scale_y_continuous(
    breaks = c(0, 150, 300, 450, 600, 750),
    labels = c("0", "150", "300", "450", "600", "750")
  ) +
  theme(axis.text.x = element_text(size = 9, vjust = 0.5, hjust = 0.5)) +
  labs(title = "Unemployment Rate vs. Number of HIV Diagnoses",
       x = "Unemployment Rate (%)",
       y = "Number of HIV Diagnoses") 

ggplotly(hiv_unemployment_plot) 
```

**_Comments_**

Overall, the smooth line indicates an increasing trend of HIV diagnoses when unemployment rate increases. 