---
title: "Model Building"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(plotly)
library(corrplot)
library(geepack)
library(fastDummies)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Analysis 
```{r, message=FALSE, warning=FALSE}
rate_complete_df = read_csv('./data/rate_complete.csv')
```

Since our outcome, HIV Diagnoses rate, is calculated using a count variable (number of HIV Diagnoses) repeated over 2013-2019, we’ll model it using Generalized Estimating Equation with a Poisson link function.

```{r}
rate_complete_df %>% 
  ggplot(aes(x = hiv_diag_rate)) + 
  geom_density() +
  theme_light() +
  labs(x = "HIV Diagnoses Rate",
       y = "Count",
       title = "Distribution of HIV Diagnoses Rate") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
rate_complete_df %>% 
  drop_na() %>%
  select(-X1, -aids_rate, -plwha_rate, -death_rate, -borough_cat, -year) %>%
  cor() %>%
  corrplot(type = "lower",
           method = "color", 
           addCoef.col = "black", 
           diag = FALSE, 
           number.cex = .55,
           insig = "blank" ,
           tl.col = "black",
           tl.cex = .9,
           tl.srt = 45
           )

```

We found that several predictors are highly correlated with each other, which should not be included in the model simultaneously, such as:

* `crime_rate` and `female` (r=0.87)
* `white` and `black` (r=-0.82)
* `american_indian` and `hawaiian` (0.98)
* `high_school` and `black` (r=-0.93)
* `age_over_64` and `black` (r=-0.97)


