---
title: "Model Building"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(plotly)
library(corrplot)
library(geepack)
library(fastDummies)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

## Analysis 
```{r, message=FALSE, warning=FALSE}
rate_complete_df = read_csv('./data/rate_complete.csv')
```

Since our outcome, HIV Diagnoses rate, is calculated using a count variable (number of HIV Diagnoses) repeated over 2013-2019, we’ll model it using Generalized Estimating Equation with a Poisson link function.

```{r, message=FALSE, warning=FALSE}
rate_complete_df %>% 
  ggplot(aes(x = hiv_diag_rate)) + 
  geom_density() +
  theme_light() +
  labs(x = "HIV Diagnoses Rate",
       y = "Density",
       title = "Distribution of HIV Diagnoses Rate") +
  theme(plot.title = element_text(hjust = 0.5))
```


<br>

## Relevant Predictors
The data we are using includes the following variables:

* `year` The year of HIV diagnosis measured.
* `hiv_diag_rate` The HIV diagnosis rate.
* `crime_rate` The crime rate.
* `unemployment_rate` The unemployment rate.
* `borough_cat` The name of borough.
* `female` The percentage of female within each borough. 
* `white` The percentage of population with Race categorized as White within each borough. 
* `black` The percentage of population with Race categorized as Black or African American within each borough. 
* `american_indian` The percentage of population with Race categorized as American Indian and Alaska Native within each borough. 
* `asian` The percentage of population with Race categorized as Asian within each borough. 
* `hawaiian` The percentage of population with Race categorized as Native Hawaiian and Other Pacific Islander within each borough. 
* `hispanic` The percentage of population with Race categorized as Hispanic or Latino within each borough. 
* `high_school` The percentage of people age 25 years+, 2014-2018, with Race categorized as High school graduate or higher degree within each borough. 
* `bachelor` The percentage of population with education categorized as Bachelor's degree or higher degree within each borough. 
* `poverty` The percentage of population in poverty within each borough. 
* `labor_force` The percentage of people age 16 years+, 2014-2018, in civilian labor force.
* `age_under_18` The percentage of population under 18 years.
* `age_over_64` The percentage of population over 64 years.
* `age_18_64` The percentage of population between 18 and 64 years.

<br>

## Correlation Matrix


Before building the model, it is important to assess crude correlation among relevant variables, in case issues of multicollinearity arise.


```{r, message=FALSE, warning=FALSE}

rate_complete_df %>% 
  drop_na() %>%
  select(-X1, -aids_rate, -plwha_rate, -death_rate, -borough_cat, -year) %>% 
  cor() %>%
  corrplot(type = "lower",
           method = "color", 
           addCoef.col = "black", 
           diag = FALSE, 
           number.cex = .45,
           insig = "blank" ,
           tl.col = "black",
           tl.cex = .7,
           tl.srt = 1
           )
```

We found that several predictors are highly correlated with each other, which should not be included in the model simultaneously, such as:

* `crime_rate` and `female` (r=0.87)
* `white` and `black` (r=-0.82)
* `american_indian` and `hawaiian` (0.98)
* `high_school` and `black` (r=-0.93)
* `age_over_64` and `black` (r=-0.97)

<br>

## Summary of Models

Based on our preliminary analysis and the correlation matrix, we propose the following models: 

1. A hypothesized model based on the effect of, `year`, `crime_rate`, `unemployment_rate`.

2. A hypothesized model based on the effect of, `year`, `black`, `asian`, `crime_rate`, `unemployment_rate`, `bachelor` and `age_18_64`.

In addition, we propose a baseline model:

3. A naive model, which only uses `year` to predict HIV diagnosis rate.

```{r, message=FALSE, warning=FALSE}
naive_model = 
  geeglm(formula = hiv_diag_rate ~ year,
         data = rate_complete_df,
         family = poisson,
         id = borough_cat,
         corstr = "ar1")

rate_df = 
  rate_complete_df %>% 
  filter(!year %in% c(2011,2012))

hyp_model =
  geeglm(hiv_diag_rate ~
         year + crime_rate + unemployment_rate,
         id = borough_cat,
         data = rate_df,
         family = poisson,
         corstr = "ar1")

hyp_model_2 =
  geeglm(hiv_diag_rate ~
         year + crime_rate + unemployment_rate + 
         black + asian + bachelor + age_18_64,
         id = borough_cat,
         data = rate_df,
         family = poisson,
         corstr = "ar1")

```


<br>

## Cross Validation


We assessed prediction accuracy of our proposed models using cross validation, which allowed us to compare the root mean squared error (RMSE) across models. Models with lower RMSEs indicates a smaller prediction error and a better prediction accuracy.

(*Note:* We split our dataset into training (80%) and testing (20%) dataset and repeated the process 100 times using the function `cross_mc`. We then used the training datasets to train our three proposed models, which were then ran on testing data. RMSEs were extracted to assess the prediction accuracy of the trained models on testing data. 100 RMSEs for each of the three models were obtained and the distributions of which were plotted using `ggplot` and `geom_violin`.)

```{r, message=FALSE, warning=FALSE}
cv_df =
  crossv_mc(rate_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(
    naive_model = map(train, ~geeglm(
        formula = hiv_diag_rate ~ borough_cat,
        family = poisson,
        id = borough_cat,
        corstr = "ar1", 
        data = .x)),
    hyp_model = map(train, ~geeglm(hiv_diag_rate ~
        borough_cat + year +
        crime_rate + unemployment_rate,
        id = borough_cat,
        family = poisson,
        corstr = "ar1", 
        data = .x)),
    hyp_model_2 = map(train, ~geeglm(hiv_diag_rate ~
        borough_cat + year,
        id = borough_cat,
        family = poisson,
        corstr = "ar1", 
        data = .x))
    ) %>% 
  mutate(
    rmse_naive = map2_dbl(naive_model, test, ~rmse(model = .x, data = .y)),
    rmse_hyp = map2_dbl(hyp_model, test, ~rmse(model = .x, data = .y)),
    rmse_hyp_2 = map2_dbl(hyp_model_2, test, ~rmse(model = .x, data = .y)))
```

<br>


```{r, message=FALSE, warning=FALSE, cache=TRUE}
cv_df %>%
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>%
  ggplot(aes(x = model, y = rmse)) +
  geom_violin(aes(fill = model)) + 
  theme_light()
  
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_") %>%
  group_by(model) %>% 
  summarize(avg_rmse = mean(rmse)) %>% 
  knitr::kable()
```


As we can see, the two hypothesized models have better predictive capabilities compared to the naive model. `hyp_model_2` has a slightly better predictive capability compared to `hyp_model`. Therefore, we select `hyp_model_2` for our further analyses.

<br>

## Conclusion

Our model can be estimated via GEE:

```{r, message=FALSE, warning=FALSE}
hyp_model_2 %>% 
  broom::tidy() %>% 
  knitr::kable()
```

According to our resulting table, all hypothesized predictors, `year`, `black`, `asian`, `crime_rate`, `unemployment_rate`, `bachelor` and `age_18_64`, are significantly associated with HIV Diagnoses rate in Bronx, Brooklyn, Manhattan, Queens, and Staten Island at the 5% level of significance. Additionally, variable `year` and boroughs with lower `unemployment_rate`, higher percentage of education (`bachelor`) and race(`asian`) are negatively associated with HIV diagnosis.

Furthermore, the corresponding interpretable estimated measure of effect can be obtained by exponentiating each parameter estimate. 

