---
title: "Model Building"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(plotly)
library(corrplot)
library(geepack)
library(fastDummies)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Analysis 
```{r, message=FALSE, warning=FALSE}
rate_complete_df = read_csv('./data/rate_complete.csv')
```

Since our outcome, HIV Diagnoses rate, is calculated using a count variable (number of HIV Diagnoses) repeated over 2013-2019, we’ll model it using Generalized Estimating Equation with a Poisson link function.

```{r}
rate_complete_df %>% 
  ggplot(aes(x = hiv_diag_rate)) + 
  geom_density() +
  theme_light() +
  labs(x = "HIV Diagnoses Rate",
       y = "Count",
       title = "Distribution of HIV Diagnoses Rate") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Relevant predictors
The data we are using includes the following variables:

* `year` The year of HIV diagnosis measured.
* `hiv_diag_rate` The HIV diagnosis rate.
* `crime_rate` The crime rate.
* `unemployment_rate` The unemployment rate.
* `borough_cat` The name of borough.
* `female` The percentage of female within each borough. 
* `white` The percentage of population with Race categorized as White within each borough. 
* `black` The percentage of population with Race categorized as Black or African American within each borough. 
* `american_indian` The percentage of population with Race categorized as American Indian and Alaska Native within each borough. 
* `asian` The percentage of population with Race categorized as Asian within each borough. 
* `hawaiian` The percentage of population with Race categorized as Native Hawaiian and Other Pacific Islander within each borough. 
* `hispanic` The percentage of population with Race categorized as Hispanic or Latino within each borough. 
* `high_school` The percentage of people age 25 years+, 2014-2018, with Race categorized as High school graduate or higher degree within each borough. 
* `bachelor` The percentage of population with education categorized as Bachelor's degree or higher degree within each borough. 
* `poverty` The percentage of population in poverty within each borough. 
* `labor_force` The percentage of people age 16 years+, 2014-2018, in civilian labor force.
* `age_under_18` The percentage of population under 18 years.
* `age_over_64` The percentage of population over 64 years.
* `age_18_64` The percentage of population between 18 and 64 years.



