---
title: "Model Building"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(plotly)
library(corrplot)
library(geepack)
library(fastDummies)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

## Analysis 
```{r, message=FALSE, warning=FALSE}
rate_complete_df = read_csv('./data/rate_complete.csv')
```

Since our outcome, HIV Diagnoses rate, is calculated using a count variable (number of HIV Diagnoses) repeated over 2013-2019, we’ll model it using Generalized Estimating Equation with a Poisson link function.

```{r, message=FALSE, warning=FALSE}
rate_complete_df %>% 
  ggplot(aes(x = hiv_diag_rate)) + 
  geom_density() +
  theme_light() +
  labs(x = "HIV Diagnoses Rate",
       y = "Count",
       title = "Distribution of HIV Diagnoses Rate") +
  theme(plot.title = element_text(hjust = 0.5))
```

<br>

## Relevant predictors
The data we are using includes the following variables:

* `year` The year of HIV diagnosis measured.
* `hiv_diag_rate` The HIV diagnosis rate.
* `crime_rate` The crime rate.
* `unemployment_rate` The unemployment rate.
* `borough_cat` The name of borough.
* `female` The percentage of female within each borough. 
* `white` The percentage of population with Race categorized as White within each borough. 
* `black` The percentage of population with Race categorized as Black or African American within each borough. 
* `american_indian` The percentage of population with Race categorized as American Indian and Alaska Native within each borough. 
* `asian` The percentage of population with Race categorized as Asian within each borough. 
* `hawaiian` The percentage of population with Race categorized as Native Hawaiian and Other Pacific Islander within each borough. 
* `hispanic` The percentage of population with Race categorized as Hispanic or Latino within each borough. 
* `high_school` The percentage of people age 25 years+, 2014-2018, with Race categorized as High school graduate or higher degree within each borough. 
* `bachelor` The percentage of population with education categorized as Bachelor's degree or higher degree within each borough. 
* `poverty` The percentage of population in poverty within each borough. 
* `labor_force` The percentage of people age 16 years+, 2014-2018, in civilian labor force.
* `age_under_18` The percentage of population under 18 years.
* `age_over_64` The percentage of population over 64 years.
* `age_18_64` The percentage of population between 18 and 64 years.

<br>

## Correlation Matrx

```{r, message=FALSE, warning=FALSE}

rate_complete_df %>% 
  drop_na() %>%
  select(-X1, -aids_rate, -plwha_rate, -death_rate, -borough_cat, -year) %>% 
  cor() %>%
  corrplot(type = "lower",
           method = "color", 
           addCoef.col = "black", 
           diag = FALSE, 
           number.cex = .45,
           insig = "blank" ,
           tl.col = "black",
           tl.cex = .7,
           tl.srt = 1
           )
```

We found that several predictors are highly correlated with each other, which should not be included in the model simultaneously, such as:

* `crime_rate` and `female` (r=0.87)
* `white` and `black` (r=-0.82)
* `american_indian` and `hawaiian` (0.98)
* `high_school` and `black` (r=-0.93)
* `age_over_64` and `black` (r=-0.97)

<br>

## Summary of Models

Based on our preliminary analysis and the correlation matrix, we propose the following models: 

1. Our hypothesized model based on the effect of `borough_cat`, `year`, `black`, `Asian`, `crime_rate`, `unemployment_rate`, `bachelor` and `age_18_64`.
2. A model selected via a GEE-specific forward stepwise selection process.

In addition, we propose a baseline model:

3. A naive model, which only uses `borough_cat` to predict HIV diagnosis rate.

```{r, message=FALSE, warning=FALSE}
naive_model = 
  geeglm(formula = hiv_diag_rate ~ year,
         data = rate_complete_df,
         family = poisson,
         id = borough_cat,
         corstr = "ar1")

rate_df = 
  rate_complete_df %>% 
  filter(!year %in% c(2011,2012))

hyp_model =
  geeglm(hiv_diag_rate ~
         year + crime_rate + unemployment_rate,
         id = borough_cat,
         data = rate_df,
         family = poisson,
         corstr = "ar1")

hyp_model_2 =
  geeglm(hiv_diag_rate ~
         year + crime_rate + unemployment_rate + 
         black + asian + bachelor + age_18_64,
         id = borough_cat,
         data = rate_df,
         family = poisson,
         corstr = "ar1")

```


<br>

## Cross Validation

```{r, message=FALSE, warning=FALSE}
cv_df =
  crossv_mc(rate_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(
    naive_model = map(train, ~geeglm(
        formula = hiv_diag_rate ~ borough_cat,
        family = poisson,
        id = borough_cat,
        corstr = "ar1", 
        data = .x)),
    hyp_model = map(train, ~geeglm(hiv_diag_rate ~
        borough_cat + year +
        crime_rate + unemployment_rate,
        id = borough_cat,
        family = poisson,
        corstr = "ar1", 
        data = .x)),
    hyp_model_2 = map(train, ~geeglm(hiv_diag_rate ~
        borough_cat + year,
        id = borough_cat,
        family = poisson,
        corstr = "ar1", 
        data = .x))
    ) %>% 
  mutate(
    rmse_naive = map2_dbl(naive_model, test, ~rmse(model = .x, data = .y)),
    rmse_hyp = map2_dbl(hyp_model, test, ~rmse(model = .x, data = .y)),
    rmse_hyp_2 = map2_dbl(hyp_model_2, test, ~rmse(model = .x, data = .y)))
```

<br>

**Compare models using cross-validated prediction error**

```{r, message=FALSE, warning=FALSE}
cv_df %>%
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>%
  ggplot(aes(x = model, y = rmse)) +
  geom_violin(aes(fill = model)) + 
  theme_light()
  
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_") %>%
  group_by(model) %>% 
  summarize(avg_rmse = mean(rmse)) %>% 
  knitr::kable()
```

<br>

## Select a Model

```{r, message=FALSE, warning=FALSE}
hyp_model_2 %>% 
  broom::tidy() %>% 
  knitr::kable()
```


