---
title: "Unemployment in NYC"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(plotly)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r, message=FALSE, warning=FALSE}
unemployment_df = read_csv('./data/unemployment_rate.csv')

unemployment_rate = 
  unemployment_df %>% 
  filter(month == "avg") %>% 
  select(unemployment_rate, area, year, month) %>% 
  mutate(
    area = as.factor(area), 
    month = as.factor(month)
  ) %>% 
  mutate(
    area = recode(area, "Bronx County" = "Bronx", "Kings County" = "Brooklyn", "New York County" = "Manhattan", "Queens County" = "Queens", "Richmond County" = "Staten Island")
    ) %>% 
  filter(area %in% c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island")) %>% 
  filter(year %in% c(2011:2019))
```


## Unemployment Distribution Over Years
```{r, message=FALSE, warning=FALSE}
unemployment_plot = 
  unemployment_rate %>% 
  ggplot(aes(x = year, y = unemployment_rate, color = area)) +
  geom_point(alpha = 0.7) +
  geom_line(size = 0.5, alpha = 0.5) +
  scale_x_continuous(
    breaks = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019),
    labels = c("2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019")
  ) +
  theme(axis.text.x = element_text(angle = 25, vjust = 0.5, hjust = 0.5)) +
  labs(title = "Unemployment Rate Within Each Borough",
       x = "Year",
       y = "Unemployment Rate (%)") 

ggplotly(unemployment_plot) %>% 
  layout(legend = list(orientation = "h", 
                       x = 0.1, y = -0.3))
```

**_Comments_**

The unemployment rate decreases over time from 2011 to 2019 for Bronx (Bronx County), Brooklyn (Kings County), Manhattan (New York County), Queens (Queens County), Staten Island (Richmond County). Specifically, there is a slight increase from 2011 to 2012, and a rapid decrease from 2012 to 2015. 




```{r, message=FALSE, warning=FALSE}
hiv_data = read_csv('./data/hiv_complete.csv')

hiv_diagnoses = 
  hiv_data %>% 
  filter(borough == "All", sex == "All") %>% 
  filter(borough_cat != "All") %>% 
  select(year, total_hiv_diag_N, borough_cat) %>% 
  filter(borough_cat %in% c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island")) %>% 
  filter(year %in% c(2011:2019)) %>% 
  rename(area = borough_cat)

hiv_unemployment = 
  merge(hiv_diagnoses, unemployment_rate)
```

## Unemployment Rate VS. HIV Diagnoses 
```{r, message=FALSE, warning=FALSE}
hiv_unemployment_plot = 
  hiv_unemployment %>% 
  mutate(
    year = as.factor(year)
  ) %>% 
  ggplot(aes(x = unemployment_rate, y = total_hiv_diag_N)) +
  geom_point(aes(shape = area, color = year)) +
  geom_smooth(size = 0.3) +
  scale_y_continuous(
    breaks = c(0, 150, 300, 450, 600, 750),
    labels = c("0", "150", "300", "450", "600", "750")
  ) +
  theme(axis.text.x = element_text(size = 9, vjust = 0.5, hjust = 0.5)) +
  labs(title = "Unemployment Rate vs. Number of HIV Diagnoses",
       x = "Unemployment Rate (%)",
       y = "Number of HIV Diagnoses") 

ggplotly(hiv_unemployment_plot) 
```

**_Comments_**

According to the plot that looks at the HIV diagnoses against unemployment rate, there are some frustration. However, overall, the smooth line indicates an increasing trend of HIV diagnoses when unemployment rate increases. 